---
title: "DPS - Project 2"
author: "Sajad, Mathias, Anders, Rasmus og Frederik"
date: "2024-04-25"
output: html_document
---

```{r}
### Polynomial Exercise 6 
library(boot)
### Regression ch 7 EX solutions 
library(splines)
### Smoothing splines: ch 7 EX solutions
library(gam)
#Detech mgcv because it interacts with gam
#detach("package:mgcv", unload = TRUE)
library('mgcv')
#Tree based methods on carseats dataset 
library(tree)
library(rpart)
# Application in R - Randomforest
library(randomForest)
# Application R - Boosting
library(gbm)
library(dplyr)
library(ggplot2)
library(ranger)
library(h2o)
# Application HOM Boosting without 
library(xgboost)
library(caret)
library(rsample)
library(foreach)
library(doParallel)
### Case study Tree - Predicting Algae blooms
library(DMwR2)
# displaying missing data
library(DataExplorer)
## QQ 
library(car)
## Boxplot distribution 
library(lattice)
# Categorial variables to displayed in a particular order
library(forcats)
# To fill NAs based on correlation
library(corrplot)
# Growing a regression tree
library(rpart.plot)
# SVM 
library(e1071)
# ROC curves - Chagning Gramma settings 
library(ROCR)
# ROC and AU with caTools
library(caTools)
# More SVM
library(kernlab)
library(pdp)
library(vip)
```


# Loading the data
```{r}
library(readr)
cardata <- read_csv("Car_data4.csv")

```

# Data preprocessing 

```{r}
library(dplyr)
library(stringr)
cardata <- cardata %>%
  mutate(Year = case_when(
    # When the format is four digits, take as is
    str_detect(`Production Year`, "^[0-9]{4}$") ~ str_extract(`Production Year`, "[0-9]{4}"),
    # When the format is two digits
    str_detect(`Production Year`, "[0-9]{2}$") ~ {
      year_part <- str_extract(`Production Year`, "[0-9]{2}")
      # Convert to four digits, assuming '20' for years 00-21 and '19' for years 22-99
      if_else(as.numeric(year_part) <= 30, str_c("20", year_part), str_c("19", year_part))
    },
    # Set to NA if the year is not in the expected format
    TRUE ~ as.character(NA)
  ))
```


## removing Cars with certain production years and unnamed columns. 

```{r}
# We remove cars that are produced in 2024 and before 1990. 
cardata <- cardata %>% filter(!(Year == 2024 | Year < 1990))
cardata$`Unnamed: 57`= NULL
cardata$`Unnamed: 58`= NULL

```

## Removing Ethanol cars

```{r}
cardata <- cardata[cardata$`Car Name` != "Renault Twingo 1,2 16V Authentique ECO2", ]

```


## Creating two new columns.
```{r}
library(tidyr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(stringr)

# Read the data (assuming the dataframe is named 'cardata')

# Separate the 'Car Name' into 'Brand' and 'Model' with additional details
cardata <- cardata %>%
  separate(`Car Name`, into = c("Brand", "Model"), sep = " ", extra = "merge")

# Function to handle Tesla brand exception
clean_model <- function(brand, model) {
  if (brand == "Tesla") {
    # Match the model to either 'Model 3', 'Model Y', 'Model S', or 'Model X'
    matched_model <- str_match(model, "(Model 3|Model Y|Model S|Model X)")[1]
    if (!is.na(matched_model)) {
      return(matched_model)
    }
  }
  # Default case for non-Tesla or when no Tesla model is matched
  return(word(model, 1))
}

# Apply the function to the 'Model' column to keep the exception for Tesla
cardata <- cardata %>%
  rowwise() %>%
  mutate(Model = clean_model(Brand, Model))

# View the updated dataframe
print(cardata)


```

## convert year to numeric type

```{r}
# Convert Year to numeric type
cardata$Year <- as.numeric(cardata$Year)
cardata$`Production Year`= NULL
```

### Seperating Motor Size and fuel to two columns for themselves
```{r}
library(dplyr)
library(stringr)

# Adjusted code to remove the first numeric part and then separate motor size and fuel type
cardata <- cardata %>%
  mutate(
    # Remove the first number (e.g., "2.5" in "2.5 Hybrid 144 kW 197 HK")
    Motor_Specs_Cleaned = str_replace(`Motor Specs`, "^[0-9]+\\.[0-9]+\\s+", ""),
    # Separate the fuel type from the remaining string
    Fuel_Type = str_extract(Motor_Specs_Cleaned, "^\\D+"),
    # Extract horsepower, which is the numeric value before "HK"
    Horsepower = str_extract(Motor_Specs_Cleaned, "\\d+\\s?(?=HK)")
  ) %>%
  # Convert extracted horsepower to numeric, handling NAs
  mutate(
    Horsepower = ifelse(is.na(Horsepower), NA_real_, as.numeric(Horsepower))
  )

# Select only the relevant columns to keep the dataframe tidy
cardata <- cardata %>%
  select(-Motor_Specs_Cleaned)

# Print out the updated dataframe to confirm changes
print(cardata)

```


### Cleaning the names for Automatgear and manuelt Gear 
```{r}
################ Automatgear and manuelt gear cleaning. 

cardata <- cardata %>%
  mutate(Gear = case_when(
    str_detect(Gear, "Automatgear") ~ "Automatgear",
    str_detect(Gear, "Manuelt gear") ~ "Manuelt gear",
    TRUE ~ as.character(NA) # Set to NA if neither pattern matches
  ))
```

### Adjustig Fuel_type and Km/L to corresponds right 

```{r}
library(dplyr)

# Update cardata by setting Km/L to NA when Fuel_Type is "EL"
cardata <- cardata %>%
  mutate(
    `Km/L` = gsub("\\n\\*", "", `Km/L`), # Remove any line break and asterisk from Km/L
    `Km/L` = ifelse(Fuel_Type == "EL", NA, as.numeric(`Km/L`)) # Set Km/L to NA for electric vehicles
  )

# Ensure that 'Km/L' is properly referenced with backticks because it contains a special character.
# Print out the updated dataframe to confirm changes
print(cardata)


```


```{r}
plot_bar(cardata, nrow = 3L, ncol = 3L)
zero_count <- sum(cardata$'KM Driven' == 0)
```


### Deleting variables - and color. 

```{r}
cardata$ABS <- NULL
cardata$Alarm <- NULL
cardata$Aircondition <- NULL
cardata$`Alloy Rim` <- NULL
cardata$Bluetooth <- NULL
cardata$Centrallock <- NULL
cardata$Elruder <- NULL
cardata$`El-spejle` <- NULL
cardata$Kurvelys <- NULL
cardata$Kørecomputer <- NULL
cardata$Multifunktionsrat <- NULL
cardata$Servo <- NULL
cardata$`Stop&Go` <- NULL
cardata$`Vejbane Advarselssystem` <- NULL
cardata$'Sport Seats' <- NULL
cardata$'Reverse Camera' <- NULL
cardata$Blindvinkelassistent <- NULL
cardata$'Digital Instrumentering' <- NULL
cardata$'Electronic Trunk Opener'  <- NULL
cardata$Color = NULL
cardata$`Electronic Trunck Opener`= NULL 
cardata$`El-Ruder`= NULL
```

### Creating a new variable based for the Range (eletric) variable 


### How many Cares have zero KL/L and deleling them 
```{r}
# Assuming 'cardata' is your dataframe and 'Km driven' is the column of interest

# Counting the zeros in the 'Km driven' column
zero_count <- sum(cardata$`KM Driven` == 0)


cardata <- cardata %>%
  filter('KM Driven' != 0)
 
```

### Setting the price to numeric and ploting the variables
```{r}
library(tidyverse)
library(DataExplorer)
cardata <- cardata %>%
  mutate(Price = as.numeric(gsub("\\.", "", Price)))
plot_histogram(cardata$Price)
plot_histogram(cardata)

```


### Lumping method to delete cars that appears less than 50 times based on the brand 

```{r}
library(dplyr)

# Assuming 'Brand' is the column that contains the car brands and 'cardata' is your dataframe.
# First, count the number of occurrences of each brand.
brand_counts <- cardata %>%
  count(Brand)

# Filter out brands that appear less than 50 times.
brands_to_keep <- brand_counts %>%
  filter(n >= 45) %>%
  select(Brand)

# Now, filter the original data to keep only those rows with brands that appear 45 times or more.
cardata <- cardata %>%
  semi_join(brands_to_keep, by = "Brand")

```


```{r}
sort(table(cardata$Fuel_Type))
library(dplyr)
library(stringr) # for string manipulation functions

# Standardize the Fuel_Type column
cardata <- cardata %>%
  mutate(
    Fuel_Type = str_trim(gsub("\\n", "", Fuel_Type)), # Remove newlines and trim white spaces
    Fuel_Type = case_when(
      str_detect(Fuel_Type, "Hybrid") ~ "Hybrid",
      str_detect(Fuel_Type, "El/Diesel") ~ "Hybrid",
      str_detect(Fuel_Type, "El/Benzin") ~ "Hybrid",
      str_detect(Fuel_Type, "El") ~ "El",
      str_detect(Fuel_Type, "Diesel") ~ "Diesel",
      str_detect(Fuel_Type, "Benzin") ~ "Benzin",
      TRUE ~ Fuel_Type # catch-all to keep other types as they are
    )
  )

# Now let's see the cleaned-up table for Fuel_Type
sort(table(cardata$Fuel_Type))

```
### Detecting outliers by brand 
```{r}

library(plotly)
library(dplyr)
# Assuming 'cardata' is your dataframe, 'Brand' is the column with car brands, and 'Price' is the column of interest

# Calculate the mean, median, 1st quartile (25th percentile), 3rd quartile (75th percentile), and 90th percentile of the 'Price' column
mean_price <- mean(cardata$Price, na.rm = TRUE)
median_price <- median(cardata$Price, na.rm = TRUE)
quantiles <- quantile(cardata$Price, probs = c(0.25, 0.75), na.rm = TRUE)
first_quartile <- quantiles[1]
third_quartile <- quantiles[2]

OutlierUpper <- third_quartile + (1.5 * (third_quartile - first_quartile))
OutlierLower <- first_quartile - (1.5 * (third_quartile - first_quartile))
cardata <- cardata %>%
  filter(Price < OutlierUpper)


# Create a scatter plot for 'Price' with 'Brand' on the x-axis
p <- plot_ly(cardata, x = ~Brand, y = ~Price, type = 'scatter', mode = 'markers',
             text = ~Brand, hoverinfo = 'text+y', marker = list(size = 5))

# Customizing layout
p <- p %>% layout(title = 'Price Outlier Detection by Brand',
                  xaxis = list(title = 'Brand', tickangle = -45),
                  yaxis = list(title = 'Price'))

# Make the plot interactive
p

```

# EDA

## Plots

```{r}
#Count the unique brands and determine number of plots
brands <- unique(cardata$Brand)
n_brands_per_plot <- 6
n_plots <- ceiling(length(brands) / n_brands_per_plot)

#Create a list to store plots
plot_list <- list()

#Create and store each plot in the list
for (i in 1:n_plots) {
  # Correcting index to make sure it does not go out of bounds
  start_index <- (i - 1) * n_brands_per_plot + 1
  end_index <- min(i * n_brands_per_plot, length(brands))
  subset_brands <- brands[start_index:end_index]
  subset_data <- cardata %>% filter(Brand %in% subset_brands)

  plot_list[[i]] <- ggplot(subset_data, aes(x = Brand, y = Price)) +
    geom_boxplot() +
    theme_minimal() +
    labs(title = paste("Boxplot of Prices for Brands", start_index, "to", end_index),
         x = "Brand",
         y = "Price") +
    scale_y_continuous(breaks = seq(100000, 600000, by = 100000), 
                       labels = c("100,000", "200,000", "300,000", "400,000", "500,000", "600,000"))
}

#Print or save plots
#For example, to print the first plot:
print(plot_list[[1]])
print(plot_list[[2]])
print(plot_list[[3]])
print(plot_list[[4]])
print(plot_list[[5]])
print(plot_list[[6]])

#To save all plots to files
for (i in seq_along(plot_list)) {
  ggsave(filename = paste("brand_boxplot", i, ".png", sep = ""), plot = plot_list[[i]], width = 10, height = 6)
}
```



```{r}
scatter_plot <- ggplot(cardata, aes(x = Year, y = Price)) +
  geom_point() +  # Adds points to the scatterplot
  theme_minimal() +  # Applies a minimal theme for aesthetic purposes
  labs(title = "Car Prices Over Years",
       x = "Year",
       y = "Price ($)")  # Labels for the plot
scatter_plot
```

### Data seperation between fuel type and range 
#### PS slet de data, som tilhøre de respektive andele for de nye data. 

```{r}
#Assuming 'cardata' is your original dataframe
#Create a new dataset for fuel types Benzin, Diesel, and Hybrid
car_data_bdh <- cardata %>%
  filter(Fuel_Type %in% c("Benzin", "Diesel", "Hybrid"))

#Create a new dataset for fuel type El (Electric)
car_data_el <- cardata %>%
  filter(Fuel_Type == "El")

#You can check the first few rows of each dataset to confirm
print(head(fuel_data_bdh))
print(head(fuel_data_el))

```



